---
title: "Reproducible code report - VANESSA-DAM-SA"
author: "Produced with [VANESSA-DAM](https://github.com/orijitghosh/VANESSA-DAM) on `r Sys.Date()`"
github: {user: orijitghosh, repo: VANESSA-DAM}
output:
  html_document:
    theme: journal
    highlight: haddock
params:
  meta: NA
  wd: NA
  modtau: NA
  ldperiod: NA
  min: NA
  light: NA
  genotype: NA
  replicate: NA
  remove: NA
  start: NA
  end: NA
  permethod: NA
  ul: NA
  ll: NA
# Reproducible code report - VANESSA-DAM-SA   
---

### All parameters used in the shiny app for analysis.
##### Metadata file name
```{r}
params$meta
```
##### Working directory (Use the directory where your metadata file and monitor files are saved)
```{r}
params$wd
```
##### Modulo-tau value for actograms in hours (generally used value = 24) 
```{r}
params$modtau
```
##### The total daylength in your experiment in hours (light + dark phase total, generally used value = 24)
```{r}
params$ldperiod
```
##### Summary time window in minutes (used in average profiles)
```{r}
params$min
```
##### Total light hours in hours (used for light-dark shading)
```{r}
params$light
```
##### Total number of monitors being analyzed
```{r}
params$genotype
```
##### If your analysis involves replicates (if no replicate used, keep value = 1)
```{r}
params$replicate
```
##### Number of days flies have to be alive to be included in analysis
```{r}
params$remove
```
##### Start day of analysis (keep 1 if you want to start from first day as mentioned in metadata file)
```{r}
params$start
```
##### End day of analysis (use value = N, where N is the last day you want to include in your analysis, e.g.: 
##### if you want to use day 1 to day 5 of data in your analysis, use start day as 1 and end day as 5)
```{r}
params$end
```

### **Set working directory of R to where your Metadata file and Monitor files are (in this case they have to be in the same directory). Replace all params$XXXX with the values mentioned above (values you have used in the shiny app).**

### Load all needed libraries:
```
library(data.table)
library(ggetho)
library(zeitgebr)
library(ggplot2)
library(ggridges)
library(readr)
library(damr)
library(sleepr)
library(behavr)
library(dplyr)
library(tidyr)
```

### Do data pre-processing:
```
setwd(params$wd)
metadata <- fread("params$meta")
    metadata <- na.omit(metadata)
    metadata_proc <- link_dam_metadata(metadata, result_dir = params$wd)
    dt <- load_dam(metadata_proc, FUN = sleepr::sleep_dam_annotation)
        dt[, moving := activity > 0]
        dt_curated <- curate_dead_animals(dt)
        dt_curated[,day:=ceiling(t/days(1))]
        dt_curated$day[dt_curated$day == 0] <- 1
        dt_curated[,phase:=ifelse(t %% hours(params$ldperiod)>hours(params$light),"Dark","Light")]
        dt_curated[,phase:=factor(phase, levels= c("Light","Dark"))]
        setbehavr(dt_curated, metadata_proc)
        lifespan_dt <- dt_curated[, .(lifespan = max(t)), by = id]
        valid_ids <- lifespan_dt[lifespan > days(params$remove), id]
        dt_curated <- dt_curated[id %in% valid_ids]
        dt_curated <- dt_curated[day %between% c(params$start, params$end)]
        setbehavr(dt_curated, metadata_proc)
        dt_curated[, uid := 1:.N, meta = T]
        dt_curated[, .(id, uid), meta = T]
        summary_dt <- (dt_curated[, .(
          sleep_fraction_all = mean(asleep),
          sleep_fraction_l = mean(asleep[phase == "Light"]),
          sleep_fraction_d = mean(asleep[phase == "Dark"])
        ), by = c("id", "day")])
        summary_dt <- dt_curated[meta=T][summary_dt]
        summary_dt_melted <- melt(summary_dt,
                                  measure.vars = patterns("sleep_fraction_"),
                                  variable.name = "phase", value.name = "sleep_fraction"
        )

        bout_dt <- bout_analysis(asleep, dt_curated)
        bout_dt[, day := ceiling(t / days(1))]
        bout_dt$day[bout_dt$day == 0] <- 1
        setbehavr(bout_dt, metadata_proc)
        bout_dt_awake <- bout_dt[asleep == FALSE]
        bout_dt <- bout_dt[asleep == TRUE, -"asleep"]
        bout_summary <- bout_dt[, .(
          latency = t[1] - ((day - 1) * 86400), first_bout_length = duration[1],
          latency_to_longest_bout = t[which.max(duration)], length_longest_bout = max(duration),
          n_bouts = .N, mean_bout_length = mean(duration), total_bout_length = sum(duration)
        ), by = c("id", "day")]
        bout_summary <- bout_dt[meta=T][bout_summary]
        bout_dt_new <- bout_dt[,phase:=ifelse(t %% hours(params$ldperiod)>hours(params$light),"Dark","Light")]
        bout_dt_new[,phase:=factor(phase, levels= c("Light","Dark"))]
        bout_summary_phase <- bout_dt_new[, .(n_bouts = .N, mean_bout_length = mean(duration),
                                              total_bout_length = sum(duration), latency = t[1] - ((day - 1) * 86400)), by = c("id", "phase", "day")
        ]
        bout_summary_phase <- bout_dt_new[meta=T][bout_summary_phase]
        bout_summary_awake <- bout_dt_awake[, .(
          latency = t[1] - ((day - 1) * 86400), first_bout_length = duration[1],
          latency_to_longest_bout = t[which.max(duration)], length_longest_bout = max(duration),
          n_bouts = .N, mean_bout_length = mean(duration), total_bout_length = sum(duration)
        ), by = c("id", "day")]
        bout_summary_awake <- bout_dt_awake[meta=T][bout_summary_awake]
        bout_dt_new_awake <- bout_dt_awake[,phase:=ifelse(t %% hours(params$ldperiod)>hours(params$light),"Dark","Light")]
        bout_dt_new_awake[,phase:=factor(phase, levels= c("Light","Dark"))]
        bout_summary_phase_awake <- bout_dt_new_awake[, .(n_bouts = .N, mean_bout_length = mean(duration),
                                                          total_bout_length = sum(duration)), by = c("id", "phase", "day")
        ]
        bout_summary_phase_awake <- bout_dt_new_awake[meta=T][bout_summary_phase_awake]
        activity_count_awake <- dt_curated[,. (
          activity_count_awake = sum(activity[asleep == FALSE])
        ), by = c("id", "day")]
        activity_count_awake <- dt_curated[meta=T][activity_count_awake]
        activity_index <- as.data.frame(bout_summary)
        activity_index$activity_index <- (activity_index$total_bout_length / 60) / activity_count_awake$activity_count_awake
```

### Theme configuration for figures:
```
       My_Theme <- theme(
      title = element_text(size = 10, face = "bold"),
      axis.title.x = element_text(size = 10, face = "bold"),
      axis.text.x = element_text(size = 8),
      axis.title.y = element_text(size = 10, face = "bold"),
      axis.text.y = element_text(size = 8),
      legend.text = element_text(size = 8),
      legend.position = "right",
      panel.background = element_rect(fill = "white"),
      panel.grid.major = element_line(
        size = 0.125,
        linetype = "solid",
        colour = "grey"
      ),
      panel.grid.minor = element_line(
        size = 0.0625,
        linetype = "solid",
        colour = "grey"
      ),
      strip.background = element_rect(fill = "white"),
      strip.text = element_text(
        colour = "black",
        size = 8,
        color = "black",
        face = "bold"
      )
    )

    My_Theme_1 <- theme(
      title = element_text(size = 6, face = "bold"),
      axis.title.x = element_text(size = 4, face = "bold"),
      axis.text.x = element_text(size = 3),
      axis.title.y = element_text(size = 3, face = "bold"),
      axis.text.y = element_text(size = 3),
      legend.text = element_text(size = 3),
      legend.position = "right",
      panel.background = element_rect(fill = "white"),
      panel.grid.major = element_line(
        size = 0.025,
        linetype = "solid",
        colour = "grey"
      ),
      panel.grid.minor = element_line(
        size = 0.0125,
        linetype = "solid",
        colour = "grey"
      ),
      strip.background = element_rect(fill = "white"),
      strip.text = element_text(
        colour = "black",
        size = 3,
        color = "black",
        face = "bold"
      )
    )

    My_Theme_2 <- theme(
      title = element_text(size = 20, face = "bold"),
      axis.title.x = element_text(size = 16, face = "bold"),
      axis.text.x = element_text(size = 14),
      axis.title.y = element_text(size = 16, face = "bold"),
      axis.text.y = element_text(size = 14),
      legend.text = element_text(size = 14),
      # aspect.ratio = 1,
      legend.position = "right",
      panel.background = element_rect(fill = "white"),
      panel.grid.major = element_line(
        size = 0.125, linetype = "solid",
        colour = "grey"
      ),
      panel.grid.minor = element_line(
        size = 0.0625, linetype = "solid",
        colour = "grey"
      )
    )

    My_Theme_3 <- theme(
      title = element_text(size = 20, face = "bold"),
      axis.title.x = element_text(size = 16, face = "bold"),
      axis.text.x = element_text(size = 14),
      axis.title.y = element_text(size = 16, face = "bold"),
      axis.text.y = element_text(size = 14),
      legend.text = element_text(size = 14),
      # aspect.ratio = .2,
      legend.position = "right",
      panel.background = element_rect(fill = "white"),
      panel.grid.major = element_line(
        size = 0.125, linetype = "solid",
        colour = "grey"
      ),
      panel.grid.minor = element_line(
        size = 0.0625, linetype = "solid",
        colour = "grey"
      )
    )
```
### Plot all ethograms:
```
p <- ggetho(dt, aes(z = asleep), summary_time_window = mins(params$min)) +
              stat_ld_annotations(height = 0.03, l_duration = hours(params$light), period = hours(params$ldperiod)) +
              stat_tile_etho() +
              scale_fill_distiller(palette = "Blues", trans = "reverse") +
              My_Theme
              ggsave(p, filename = "Plot all ethograms.png", type = "cairo",height = 24,
       width = 15, dpi = 600, units = "in")    ###############height = 8*n where n is number of monitors
```
### Plot curated ethograms:
```
p <- ggetho(dt_curated, aes(z = asleep), summary_time_window = mins(params$min)) +
            ggetho(dt_curated, aes(z = asleep), summary_time_window = mins(params$min)) +
              stat_ld_annotations(height = 0.03, l_duration = hours(params$light), period = hours(params$ldperiod)) +
              stat_tile_etho() +
              scale_fill_distiller(palette = "Blues", trans = "reverse") +
              My_Theme
              ggsave(p, filename = "Plot curated ethograms.png", type = "cairo",height = 24,
       width = 15, dpi = 600, units = "in")    ###############height = 8*n where n is number of monitors
```
### Plot curated ethograms wrapped:
```
p <- ggetho(dt_curated, aes(z = asleep), time_wrap = hours(params$modtau), summary_time_window = mins(params$min)) +
            ggetho(dt_curated, aes(z = asleep), time_wrap = hours(params$modtau), summary_time_window = mins(params$min)) +
              stat_ld_annotations(height = 0.03, l_duration = hours(params$light), period = hours(params$ldperiod)) +
              stat_tile_etho() +
              scale_fill_distiller(palette = "Blues", trans = "reverse") +
              My_Theme
              ggsave(p, filename = "Plot curated ethograms wrapped.png", type = "cairo",height = 24,
       width = 7, dpi = 600, units = "in")    ###############height = 8*n where n is number of monitors
```
### Plot plots over days individual:
```
p <- ggetho(dt_curated, aes(y = asleep, colour = genotype), summary_time_window = mins(params$min)) +
              stat_pop_etho() +
              stat_ld_annotations(height = 1, alpha = 0.1, outline = NA, l_duration = hours(params$light), period = hours(params$ldperiod)) +
              My_Theme +
              facet_wrap(~ genotype + id + replicate, ncol = 1, scales = "free_y")
              ggsave(p, filename = "Plot verage plots over days individual.png", type = "cairo",height = 12,
       width = 3, dpi = 600, units = "in")    ###############height = 4*n where n is number of monitors
```
### Plot Average plot wrapped individual:
```
p <- ggetho(dt_curated, aes(y = asleep, colour = genotype),
              time_wrap = hours(params$modtau), summary_time_window = mins(params$min)
            ) +
              stat_pop_etho() +
              stat_ld_annotations(height = 1, alpha = 0.1, outline = NA, l_duration = hours(params$light), period = hours(params$ldperiod)) +
              My_Theme +
              facet_wrap(~ genotype + id + replicate, ncol = 4, scales = "free_y")
              ggsave(p, filename = "Plot Average plot wrapped individual.png", type = "cairo",height = 15,
       width = 16, dpi = 600, units = "in")    ###############height = 5*n where n is number of monitors
```
### Plot Average plots over days:
```
p <- ggetho(dt_curated, aes(y = asleep, colour = genotype), summary_time_window = mins(params$min)) +
              stat_pop_etho() +
              stat_ld_annotations(height = 1, alpha = 0.1, outline = NA, l_duration = hours(params$light), period = hours(params$ldperiod)) +
              My_Theme +
              facet_grid(~replicate,  scales = "free_y")
              ggsave(p, filename = "Plot Plot Average plots over days.png", type = "cairo",height = 3,
       width = 12, dpi = 600, units = "in")    ###############height = 1*n where n is number of monitors
```
### Plot Average plot wrapped:
```
p <- ggetho(dt_curated, aes(y = asleep, colour = genotype),
              time_wrap = hours(params$modtau), summary_time_window = mins(params$min)
            ) +
              stat_pop_etho() +
              stat_ld_annotations(height = 1, alpha = 0.1, outline = NA, l_duration = hours(params$light), period = hours(params$ldperiod)) +
              My_Theme +
              facet_grid(~replicate,  scales = "free_y")
              ggsave(p, filename = "Plot Average plot wrapped.png", type = "cairo",height = 3,
       width = 10, dpi = 600, units = "in")    ###############height = 1*n where n is number of monitors
```
### Plot Average plot wrapped polar:
```
p <- ggetho(dt_curated, aes(y = asleep, colour = genotype),
              time_wrap = hours(params$modtau), summary_time_window = mins(params$min)
            ) +
              stat_pop_etho(geom = "bar", alpha = 0.4) +
              stat_ld_annotations(height = 1, alpha = .1, x_limits = c(0, hours(24)), outline = NA, l_duration = hours(params$light), period = hours(params$ldperiod)) +
              scale_y_continuous(name = "Fraction of time sleeping") +
              My_Theme +
              facet_grid(~replicate) +
              coord_polar(clip = "off")
              ggsave(p, filename = "Plot Average plot wrapped polar.png", type = "cairo",height = 7,
       width = 7, dpi = 600, units = "in")    ###############height = 1*n where n is number of monitors
```
### Plot Sleep fraction:
```
p <- ggplot(summary_dt, aes(x = genotype, y = sleep_fraction_all, fill = genotype)) +
              # geom_boxplot(outlier.colour = "red") +    ######if boxplot needed
              geom_jitter(aes(colour = genotype), size = 2,  alpha = .5, position = position_jitter(height = .1, width = .1)) +
              geom_violin(aes(color = genotype, fill = genotype), trim = TRUE, alpha = 0.5) +
              stat_summary(fun = mean, geom = "point", aes(color = genotype), size = 3, shape = 23) +
              stat_summary(aes(label = round((..y..), 2), color = genotype),
                fun = mean, geom = "text",
                size = 5, vjust = -0.5
              ) +
              scale_y_continuous(name = "Fraction of time sleeping") +
              My_Theme_2 +
              facet_grid(day~replicate,  scales = "free_y")
              ggsave(p, filename = "Plot Sleep fraction.png", type = "cairo",height = 7,
       width = 10, dpi = 600, units = "in")    ###############height = 1*n where n is number of monitors
```
### Plot Sleep fraction summary:
```
p <- ggplot(summary_dt_melted, aes(x = phase, y = sleep_fraction, fill = genotype)) +
              # geom_boxplot(outlier.colour = "red") +  ##########if boxplot needed
              geom_jitter(aes(colour = genotype), alpha = .5, position = position_dodge(.9)) +
              geom_violin(aes(color = genotype, fill = genotype), trim = TRUE, alpha = 0.5) +
              stat_summary(
                fun = mean, geom = "point", aes(color = genotype, group = genotype),
                position = position_dodge(.9), size = 3, shape = 23
              ) +
              stat_summary(aes(label = round((..y..), 2), color = genotype, group = genotype),
                fun = mean, geom = "text",
                size = 5, vjust = -0.5, position = position_dodge(.9)
              ) +
              scale_y_continuous(name = "Fraction of time sleeping") +
              My_Theme_2 +
              facet_grid(day~replicate,  scales = "free_y")
              ggsave(p, filename = "Plot Sleep fraction summary.png", type = "cairo",height = 7,
       width = 10, dpi = 600, units = "in")    ###############height = 1*n where n is number of monitors
```
### Plot Time spent sleeping:
```
p <- ggplot(bout_summary, aes(x = genotype, y = total_bout_length / 60, fill = genotype)) +
        # geom_boxplot(outlier.colour = "red") +  #######if boxplot needed
        geom_jitter(aes(colour = genotype), alpha = .5, position = position_dodge(.9)) +
        geom_violin(aes(color = genotype, fill = genotype), trim = TRUE, alpha = 0.5) +
        stat_summary(
          fun = mean, geom = "point", aes(color = genotype, group = genotype),
          position = position_dodge(.9), size = 3, shape = 23
        ) +
        stat_summary(aes(label = round((..y..), 2), color = genotype, group = genotype),
                     fun = mean, geom = "text",
                     size = 5, vjust = -0.5, position = position_dodge(.9)
        ) +
        scale_x_discrete(name = "Genotype") +
        scale_y_continuous(name = "Total time sleeping (m)") +
          facet_grid(day~replicate,  scales = "free_y") +
          My_Theme_2
              ggsave(p, filename = "Plot Time spent sleeping.png", type = "cairo",height = 7,
       width = 10, dpi = 600, units = "in")    ###############height = 1*n where n is number of monitors
```
### Plot Time spent awake:
```
p <- ggplot(bout_summary_awake, aes(x = genotype, y = total_bout_length / 60, fill = genotype)) +
              # geom_boxplot(outlier.colour = "red") +  #######if boxplot needed
              geom_jitter(aes(colour = genotype), alpha = .5, position = position_dodge(.9)) +
              geom_violin(aes(color = genotype, fill = genotype), trim = TRUE, alpha = 0.5) +
              stat_summary(
                fun = mean, geom = "point", aes(color = genotype, group = genotype),
                position = position_dodge(.9), size = 3, shape = 23
              ) +
              stat_summary(aes(label = round((..y..), 2), color = genotype, group = genotype),
                           fun = mean, geom = "text",
                           size = 5, vjust = -0.5, position = position_dodge(.9)
              ) +
              scale_x_discrete(name = "Genotype") +
              scale_y_continuous(name = "Total time awake (m)") +
              facet_grid(day~replicate,  scales = "free_y") +
              My_Theme_2
              ggsave(p, filename = "Plot Time spent awak.png", type = "cairo",height = 7,
       width = 10, dpi = 600, units = "in")    ###############height = 1*n where n is number of monitors
```
### Plot Time spent sleeping in day and night:
```
p <- ggplot(bout_summary_phase, aes(x = phase, y = total_bout_length / 60, fill = genotype)) +
              # geom_boxplot(outlier.colour = "red") +  #######if boxplot needed
              geom_jitter(aes(colour = genotype), alpha = .5, position = position_dodge(.9)) +
              geom_violin(aes(color = genotype, fill = genotype), trim = TRUE, alpha = 0.5) +
              stat_summary(
                fun = mean, geom = "point", aes(color = genotype, group = genotype),
                position = position_dodge(.9), size = 3, shape = 23
              ) +
              stat_summary(aes(label = round((..y..), 2), color = genotype, group = genotype),
                           fun = mean, geom = "text",
                           size = 5, vjust = -0.5, position = position_dodge(.9)
              ) +
              scale_x_discrete(name = "Genotype") +
              scale_y_continuous(name = "Total time sleeping (m)") +
              facet_grid(day~replicate,  scales = "free_y") +
              My_Theme_2
              ggsave(p, filename = "Plot Time spent sleeping in day and night.png", type = "cairo",height = 7,
       width = 10, dpi = 600, units = "in")    ###############height = 1*n where n is number of monitors
```
### Plot Time spent awake in day and night:
```
p <- ggplot(bout_summary_phase, aes(x = phase, y = total_bout_length / 60, fill = genotype)) +
              # geom_boxplot(outlier.colour = "red") +  #######if boxplot needed
              geom_jitter(aes(colour = genotype), alpha = .5, position = position_dodge(.9)) +
              geom_violin(aes(color = genotype, fill = genotype), trim = TRUE, alpha = 0.5) +
              stat_summary(
                fun = mean, geom = "point", aes(color = genotype, group = genotype),
                position = position_dodge(.9), size = 3, shape = 23
              ) +
              stat_summary(aes(label = round((..y..), 2), color = genotype, group = genotype),
                           fun = mean, geom = "text",
                           size = 5, vjust = -0.5, position = position_dodge(.9)
              ) +
              scale_x_discrete(name = "Genotype") +
              scale_y_continuous(name = "Total time sleeping (m)") +
              facet_grid(day~replicate,  scales = "free_y") +
              My_Theme_2
              ggsave(p, filename = "Plot Time spent awake in day and night.png", type = "cairo",height = 7,
       width = 10, dpi = 600, units = "in")    ###############height = 1*n where n is number of monitors
```
### Plot Activity Index:
```
p <- ggplot(activity_index, aes(x = genotype, y = activity_index, fill = genotype)) +
        # geom_boxplot(outlier.colour = "red") +    ######if boxplot needed
        geom_jitter(aes(colour = genotype), size = 2, alpha = .5, position = position_jitter(height = .1, width = .1)) +
        geom_violin(aes(color = genotype, fill = genotype), trim = TRUE, alpha = 0.5) +
        stat_summary(fun = mean, geom = "point", aes(color = genotype), size = 3, shape = 23) +
        stat_summary(aes(label = round((..y..), 2), color = genotype),
                     fun = mean, geom = "text",
                     size = 5, vjust = -0.5
        ) +
        scale_y_continuous(name = "Activity index") +
          My_Theme_2 +
        facet_grid(day~replicate, scales = "free_y")
              ggsave(p, filename = "Plot Activity Index.png", type = "cairo",height = 7,
       width = 10, dpi = 600, units = "in")    ###############height = 1*n where n is number of monitors
```
### Plot Bouts:
```
p <- ggetho(bout_dt, aes(y = duration / 60, colour = genotype), time_wrap = hours(params$modtau), summary_time_window = mins(params$min)) +
              stat_pop_etho() +
              stat_ld_annotations(height = 1, alpha = 0.1, outline = NA, l_duration = hours(params$light), period = hours(params$ldperiod)) +
              scale_y_continuous(name = "Bout length (min)") +
              My_Theme_3 +
              facet_grid(~replicate,  scales = "free_y")
              ggsave(p, filename = "Plot Bouts.png", type = "cairo",height = 3,
       width = 10, dpi = 600, units = "in")    ###############height = 1*n where n is number of monitors
```
### Plot Number of bouts:
```
p <- ggplot(bout_summary, aes(genotype, n_bouts, colour = genotype)) +
              geom_violin(aes(color = genotype, fill = genotype), trim = TRUE, alpha = 0.5) +
              geom_jitter(aes(colour = genotype), alpha = .5, position = position_jitter(height = .2, width = .2)) +
              stat_summary(
                fun = mean, geom = "point", aes(color = genotype, group = genotype),
                position = position_dodge(.9), size = 3, shape = 23
              ) +
              stat_summary(aes(label = round((..y..), 2), color = genotype, group = genotype),
                fun = mean, geom = "text",
                size = 5, vjust = -0.5, position = position_dodge(.9)
              ) +
              scale_x_discrete(name = "Genotype") +
              scale_y_continuous(name = "Number of bouts") +
              facet_grid(day~replicate,  scales = "free_y") +
              My_Theme_2
              ggsave(p, filename = "Plot Number of bouts.png", type = "cairo",height = 5,
       width = 7, dpi = 600, units = "in")    ###############height = 1*n where n is number of monitors
```
### Plot Number of awake bouts:
```
p <- ggplot(bout_summary_awake, aes(genotype, n_bouts, colour = genotype)) +
              geom_violin(aes(color = genotype, fill = genotype), trim = TRUE, alpha = 0.5) +
              geom_jitter(aes(colour = genotype), alpha = .5, position = position_jitter(height = .2, width = .2)) +
              stat_summary(
                fun = mean, geom = "point", aes(color = genotype, group = genotype),
                position = position_dodge(.9), size = 3, shape = 23
              ) +
              stat_summary(aes(label = round((..y..), 2), color = genotype, group = genotype),
                           fun = mean, geom = "text",
                           size = 5, vjust = -0.5, position = position_dodge(.9)
              ) +
              scale_x_discrete(name = "Genotype") +
              scale_y_continuous(name = "Number of awake bouts") +
              facet_grid(day~replicate,  scales = "free_y") +
              My_Theme_2
              ggsave(p, filename = "Plot Number of awake bouts.png", type = "cairo",height = 5,
       width = 7, dpi = 600, units = "in")    ###############height = 1*n where n is number of monitors
```
### Plot Mean bout length:
```
p <- ggplot(bout_summary, aes(genotype, mean_bout_length / 60, colour = genotype)) +
              geom_violin(aes(color = genotype, fill = genotype), trim = TRUE, alpha = 0.5) +
              geom_jitter(aes(colour = genotype), alpha = .5, position = position_jitter(height = .2, width = .2)) +
              stat_summary(
                fun = mean, geom = "point", aes(color = genotype, group = genotype),
                position = position_dodge(.9), size = 3, shape = 23
              ) +
              stat_summary(aes(label = round((..y..), 2), color = genotype, group = genotype),
                fun = mean, geom = "text",
                size = 5, vjust = -0.5, position = position_dodge(.9)
              ) +
              scale_x_discrete(name = "Genotype") +
              scale_y_continuous(name = "Mean bout length (m)") +
              facet_grid(day~replicate,  scales = "free_y") +
              My_Theme_2
              ggsave(p, filename = "Plot Mean bout length.png", type = "cairo",height = 5,
       width = 7, dpi = 600, units = "in")    ###############height = 1*n where n is number of monitors
```
### Plot Mean awake bout length:
```
p <- ggplot(bout_summary_awake, aes(genotype, mean_bout_length / 60, colour = genotype)) +
              geom_violin(aes(color = genotype, fill = genotype), trim = TRUE, alpha = 0.5) +
              geom_jitter(aes(colour = genotype), alpha = .5, position = position_jitter(height = .2, width = .2)) +
              stat_summary(
                fun = mean, geom = "point", aes(color = genotype, group = genotype),
                position = position_dodge(.9), size = 3, shape = 23
              ) +
              stat_summary(aes(label = round((..y..), 2), color = genotype, group = genotype),
                           fun = mean, geom = "text",
                           size = 5, vjust = -0.5, position = position_dodge(.9)
              ) +
              scale_x_discrete(name = "Genotype") +
              scale_y_continuous(name = "Mean awake bout length (m)") +
              facet_grid(day~replicate,  scales = "free_y") +
              My_Theme_2
              ggsave(p, filename = "Plot Mean awake bout length.png", type = "cairo",height = 5,
       width = 7, dpi = 600, units = "in")    ###############height = 1*n where n is number of monitors
```
### Plot Number of bouts in light and dark phase:
```
p <- ggplot(bout_summary_phase, aes(x = phase, y = n_bouts, fill = genotype)) +
              # geom_boxplot(outlier.colour = "red") +   #########if boxplots needed
              geom_jitter(aes(colour = genotype), alpha = .5, position = position_dodge(.9)) +
              geom_violin(aes(color = genotype, fill = genotype), trim = TRUE, alpha = 0.5) +
              stat_summary(
                fun = mean, geom = "point", aes(color = genotype, group = genotype),
                position = position_dodge(.9), size = 3, shape = 23
              ) +
              stat_summary(aes(label = round((..y..), 2), color = genotype, group = genotype),
                fun = mean, geom = "text",
                size = 5, vjust = -0.5, position = position_dodge(.9)
              ) +
              scale_x_discrete(name = "Genotype") +
              scale_y_continuous(name = "Number of bouts") +
              facet_grid(day~replicate,  scales = "free_y") +
              My_Theme_2
              ggsave(p, filename = "Plot Number of bouts in light and dark phase.png", type = "cairo",height = 5,
       width = 7, dpi = 600, units = "in")    ###############height = 1*n where n is number of monitors
```
### Plot Mean bout length in light and dark phase:
```
p <- ggplot(bout_summary_phase, aes(x = phase, y = mean_bout_length / 60, fill = genotype)) +
              # geom_boxplot(outlier.colour = "red") +  #######if boxplot needed
              geom_jitter(aes(colour = genotype), alpha = .5, position = position_dodge(.9)) +
              geom_violin(aes(color = genotype, fill = genotype), trim = TRUE, alpha = 0.5) +
              stat_summary(
                fun = mean, geom = "point", aes(color = genotype, group = genotype),
                position = position_dodge(.9), size = 3, shape = 23
              ) +
              stat_summary(aes(label = round((..y..), 2), color = genotype, group = genotype),
                fun = mean, geom = "text",
                size = 5, vjust = -0.5, position = position_dodge(.9)
              ) +
              scale_x_discrete(name = "Genotype") +
              scale_y_continuous(name = "Mean bout length (m)") +
              facet_grid(day~replicate,  scales = "free_y") +
              My_Theme_2
              ggsave(p, filename = "Plot  Mean bout length in light and dark phase.png", type = "cairo",height = 5,
       width = 7, dpi = 600, units = "in")    ###############height = 1*n where n is number of monitors
```
### Plot Mean bout length distribution:
```
p <- ggplot(bout_summary, aes((mean_bout_length / 60), genotype, color = genotype, fill = genotype)) +
              geom_density_ridges(
                scale = 1, rel_min_height = 0.01, jittered_points = TRUE,
                position = position_points_jitter(width = 0.2, height = 0),
                point_shape = "|", point_size = 8, point_alpha = 1, alpha = .7
              ) +
              facet_wrap(day~replicate, ncol = 1) +
              scale_y_discrete(expand = c(0, 0)) + # will generally have to set the `expand` option
              scale_x_continuous(expand = c(0, 0)) + # for both axes to remove unneeded padding
              coord_cartesian(clip = "off") + # to avoid clipping of the very top of the top ridgeline
              labs(x = "Mean bout length (m)", y = "Genotype") +
              facet_grid(day~replicate,  scales = "free_y") +
              My_Theme_2
              ggsave(p, filename = "Plot Mean bout length distribution.png", type = "cairo",height = 5,
       width = 7, dpi = 600, units = "in")    ###############height = 1*n where n is number of monitors
```
### Plot Latency to first bout:
```
p <- ggplot(bout_summary, aes(genotype, latency /  60, colour = genotype)) +
              geom_violin(aes(color = genotype, fill = genotype), trim = TRUE, alpha = 0.5) +
              geom_jitter(aes(colour = genotype), alpha = .5, position = position_jitter(height = .2, width = .2)) +
              stat_summary(
                fun = mean, geom = "point", aes(color = genotype, group = genotype),
                position = position_dodge(.9), size = 3, shape = 23
              ) +
              stat_summary(aes(label = round((..y..), 2), color = genotype, group = genotype),
                           fun = mean, geom = "text",
                           size = 5, vjust = -0.5, position = position_dodge(.9)
              ) +
              scale_x_discrete(name = "Genotype") +
              scale_y_continuous(name = "Latency (min)") +
              facet_grid(day~replicate,  scales = "free_y") +
              My_Theme_2
              ggsave(p, filename = "Plot Latency to first bout.png", type = "cairo",height = 5,
       width = 7, dpi = 600, units = "in")    ###############height = 1*n where n is number of monitors
```
### Plot Latency to first bout in light and dark:
```
p <- ggplot(bout_summary_phase, aes(phase, latency /  60, colour = genotype)) +
              geom_jitter(aes(colour = genotype), alpha = .5, position = position_dodge(.9)) +
              geom_violin(aes(color = genotype, fill = genotype), trim = TRUE, alpha = 0.5) +
              stat_summary(
                fun = mean, geom = "point", aes(color = genotype, group = genotype),
                position = position_dodge(.9), size = 3, shape = 23
              ) +
              stat_summary(aes(label = round((..y..), 2), color = genotype, group = genotype),
                           fun = mean, geom = "text",
                           size = 5, vjust = -0.5, position = position_dodge(.9)
              ) +
              scale_x_discrete(name = "Phase") +
              scale_y_continuous(name = "Latency (min)") +
              facet_grid(day~replicate,  scales = "free_y") +
              My_Theme_2
              ggsave(p, filename = "Plot Latency to first bout in light and dark.png", type = "cairo",height = 5,
       width = 7, dpi = 600, units = "in")    ###############height = 1*n where n is number of monitors
```
### Plot Bout summary:
```
p <- ggplot(bout_summary, aes(n_bouts, mean_bout_length / 60, colour = genotype)) +
              geom_point(size = 4, alpha = 0.5) +
              scale_x_continuous(name = "Number of bouts") +
              scale_y_continuous(name = "Average bout duration (m)") +
              facet_grid(day~replicate,  scales = "free_y") +
              My_Theme_2
              ggsave(p, filename = "Plot Bout summary.png", type = "cairo",height = 5,
       width = 10, dpi = 600, units = "in")    ###############height = 1*n where n is number of monitors
```
### Write summary tables:
```
pro_chi_sq <- bout_summary
      pro_chi_sq_new <- bout_summary_phase
      pro_chi_sq$latency <- pro_chi_sq$latency / 60
      pro_chi_sq$first_bout_length <- pro_chi_sq$first_bout_length / 60
      pro_chi_sq$latency_to_longest_bout <- pro_chi_sq$latency_to_longest_bout / 60
      pro_chi_sq$length_longest_bout <- pro_chi_sq$length_longest_bout / 60
      pro_chi_sq$mean_bout_length <- pro_chi_sq$mean_bout_length / 60
      pro_chi_sq$mean_bout_length <- pro_chi_sq$total_bout_length / 60
      pro_chi_sq <- as.matrix(pro_chi_sq)
      pro_chi_sq_new$mean_bout_length <- pro_chi_sq_new$mean_bout_length / 60
      pro_chi_sq_new$mean_bout_length <- pro_chi_sq_new$total_bout_length / 60
      pro_chi_sq_new <- as.matrix(pro_chi_sq_new)
      popplot <- ggetho(dt_curated, aes(y = asleep, colour = genotype), summary_time_window = mins(15)) +
        stat_pop_etho()
      popplot_wrap <- ggetho(dt_curated, aes(y = asleep, colour = genotype), summary_time_window = mins(15),
                             time_wrap = hours(24)) +
        stat_pop_etho()
      df_new <- ggplot_build(popplot)$plot$data   ##########works perfectly - individuals with each day profile
      df_new <- df_new[, -c("file_info")]
      # df_new <- as.data.table(df_new)
      df_new_wrap <- ggplot_build(popplot_wrap)$plot$data  ###individuals wrapped over days
      df_new_wrap_genotype <- as.data.table(df_new_wrap)
      df_new_wrap_genotype_summary_replicate <- (df_new_wrap_genotype[, .(
        sleep_fraction_genotype = mean(asleep)
      ), by = c("genotype", "t", "replicate")])   ##########average profile over genotype replicates separate
      write.csv(df_new, file = "individual sleep profiles.csv", row.names = FALSE)
      write.csv(df_new_wrap_genotype_summary_replicate, file = "average sleep profiles.csv", row.names = FALSE)
      write.csv(pro_chi_sq, file = "bout analysis data.csv", row.names = FALSE)
      write.csv(pro_chi_sq_new, file = "bout analysis data with phase details.csv", row.names = FALSE)
      
```
